---
title: "EDA"
output:
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}
proj_dir = rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir = proj_dir)
setwd("/cloud/project/02 EDA")
knitr::opts_chunk$set(echo=FALSE)
```

```{r, echo = FALSE}
library(leaflet)
library(polyCub)
library(rgdal)
library(htmlwidgets)
library(gridExtra)
library(ggpubr)
library(corrplot)
library(RColorBrewer)
source("https://raw.githubusercontent.com/janhove/janhove.github.io/master/RCode/sortLvls.R")
```

```{r load_envr, include=FALSE}
source("00 source.R")
load("00 envr/Compulsory/policy_claims.Rda")
load("00 envr/Compulsory/ass_rfct.Rda")
australian_postcodes <- read_csv("00 envr/australian_postcodes.csv")

options(scipen=999)
theme_set(theme_light())
```

```{r include=FALSE}
col1 = "#4B2B74"
col2 = "#745C94"
col3 = "#AD84C6"
col4 = "#6997AF"
col5 = "#84ACB6"
col6 = "#6F8183"

colpal = colorRampPalette(c(col1, col2, col3, col4, col5, col6))

```



```{r include=FALSE}
policy_claims <- policy_claims %>%
  replace_with_na(replace = list(grossincurred_prop = 0,
                             grossincurred_lossofinc = 0)) %>%
  mutate(
    log_grossincurred_prop = log(grossincurred_prop), 
    log_grossincurred_lossofinc = log(grossincurred_lossofinc),
    frequency_prop = replace_na(claimcount_prop / exposure, 0),
    frequency_bi = replace_na(claimcount_lossofinc / exposure, 0),
    severity_prop = replace_na(grossincurred_prop / claimcount_prop, 0),
    severity_bi = replace_na(grossincurred_lossofinc / claimcount_lossofinc, 0),
    log_severity_prop = replace_na(log_grossincurred_prop / claimcount_prop, 0),
    log_severity_bi = replace_na(grossincurred_lossofinc/ claimcount_lossofinc, 0)
  )
```

# Missing variables

The data provided had a large number of missing values, but all of them were either due to not having a loss of income coverage, or that details about the building’s characteristics were simply not collected. Setting them the values as unknowns sufficed to resolve the issue.

```{r }
###png(filename="02 EDA/missing.##png", width = 800, height = 600)
gg_miss_upset(ass_rfct, nsets=7, nintersects=NA)
###dev.off
```

# Geographic Location

```{r include=FALSE}
# add postcode back in
ass_rfct_postcode <- ass_rfct %>%
  dplyr::select(policyno, situation_num, riskpostcode) %>%
  dplyr::group_by(policyno, situation_num) %>%
  filter(row_number() == n())

policy_claims_inc_postcode <- left_join(policy_claims, ass_rfct_postcode, by = c("policyno" = "policyno", "situation_num" = "situation_num"))

sa4_postcodes <-australian_postcodes %>%
  dplyr:: select(c("postcode", "sa4name", "sa4", "Lat_precise", "Long_precise")) %>%
  mutate(postcode = as.numeric(as.character(postcode))) %>%
  group_by(postcode) %>%
  arrange(sa4name) %>%
  filter(row_number() == 1)
policy_claims_inc_postcode$riskpostcode <- as.numeric(as.character(policy_claims_inc_postcode$riskpostcode))

policy_claims_sa4 <- left_join(policy_claims_inc_postcode, sa4_postcodes, by = c("riskpostcode" = "postcode")) 

by_state <- policy_claims_sa4 %>%
  group_by(state) %>%
  summarise(
    n_policies = n(),
    lat = mean(Lat_precise),
    long = mean(Long_precise),
    mean_prop_SI = mean(suminsured_prop, na.rm = TRUE),
    mean_lossofinc_SI = mean(suminsured_lossofinc, na.rm = TRUE),
    mean_prop_grossincurred = mean(log_grossincurred_prop, na.rm = TRUE),
    mean_lossofinc_grossincurred = mean(log_grossincurred_prop, na.rm = TRUE),
    mean_prop_claims = mean(claimcount_prop, na.rm = TRUE),
    mean_lossofinc_claims = mean(claimcount_lossofinc, na.rm = TRUE)
  )
by_sa4 <- policy_claims_sa4 %>%
  group_by(sa4name, sa4) %>%
  summarise(
    n_policies = n(),
    lat = mean(Lat_precise),
    long = mean(Long_precise),
    mean_prop_SI = mean(suminsured_prop, na.rm = TRUE),
    mean_lossofinc_SI = mean(suminsured_lossofinc, na.rm = TRUE),
    mean_prop_grossincurred = mean(log_grossincurred_prop, na.rm = TRUE),
    mean_lossofinc_grossincurred = mean(log_grossincurred_prop, na.rm = TRUE),
    mean_prop_claims = mean(claimcount_prop, na.rm = TRUE),
    mean_lossofinc_claims = mean(claimcount_lossofinc, na.rm = TRUE)
  )

by_postcode <- policy_claims_sa4 %>%
  group_by(riskpostcode) %>%
  summarise(
    n_policies = n(),
    lat = mean(Lat_precise),
    long = mean(Long_precise),
    mean_prop_SI = mean(suminsured_prop, na.rm = TRUE),
    mean_lossofinc_SI = mean(suminsured_lossofinc, na.rm = TRUE),
    mean_prop_grossincurred = mean(log_grossincurred_prop, na.rm = TRUE),
    mean_lossofinc_grossincurred = mean(log_grossincurred_prop, na.rm = TRUE),
    mean_prop_claims = mean(claimcount_prop, na.rm = TRUE),
    mean_lossofinc_claims = mean(claimcount_lossofinc, na.rm = TRUE)
  )

```

## Map by SA4

By using geocode, can more easily observe trends such as concentration of claims around CBD regions, particularly in Sydney and ACT whilst maintaining the granularity that can identify differences of different regions within a state. For example, in NSW, majority of claims occur in Sydney region and increases as you move away from CBD

```{r}
# prepare colour code
pal <- colorNumeric(
  palette = "BuPu", 
  domain = by_sa4$mean_prop_grossincurred,
  na.color = NA)

# prepare tooltip
mytext <- paste(
   "SA4: ", by_sa4$sa4name, "<br/>", 
   "Number of policies: ", by_sa4$n_policies, "<br/>", 
   "Mean Number of Property claims: ", by_sa4$mean_prop_claims, "<br/>", 
   "Mean Log Gross Property Gross Incurred: ", by_sa4$mean_prop_grossincurred, sep="") %>%
  lapply(htmltools::HTML)

# create map
m <- leaflet(data = by_sa4) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lat=-27.833, lng=120 , zoom=4) %>%
  addCircleMarkers(
    lng = ~ long,
    lat = ~ lat,
    radius = ~mean_prop_claims*180,
    color = ~pal(mean_prop_grossincurred),
    weight = 1,
    popup = mytext,
    opacity = 0.8
  ) %>%
  addLegend("bottomleft",
            pal = pal, values = ~ mean_prop_grossincurred,
            title = "Mean Property Log Gross Incurred")

saveWidget(m, file="02 EDA/01 Map By SA4/PropertyClaimsSA4.html")
m
```

```{r}
# prepare colour code
pal <- colorNumeric(
  palette = "BuPu", 
  domain = by_sa4$mean_lossofinc_grossincurred,
  na.color = NA)

# prepare tooltip
mytext <- paste(
   "SA4: ", by_sa4$sa4name, "<br/>", 
   "Number of policies: ", by_sa4$n_policies, "<br/>", 
   "Mean Number of Loss Of Income claims: ", by_sa4$mean_lossofinc_claims, "<br/>", 
   "Mean Log Gross Loss Of Income Gross Incurred: ", by_sa4$mean_lossofinc_grossincurred, sep="") %>%
  lapply(htmltools::HTML)

# create map
m <- leaflet(data = by_sa4) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lat=-27.833, lng=120 , zoom=4) %>%
  addCircleMarkers(
    lng = ~ long,
    lat = ~ lat,
    radius = ~mean_lossofinc_claims*2000,
    color = ~pal(mean_lossofinc_grossincurred),
    weight = 1,
    popup = mytext,
    opacity = 0.8
  ) %>%
  addLegend("bottomleft",
            pal = pal, values = ~ mean_lossofinc_grossincurred,
            title = "Mean Loss Of Income Log Gross Incurred")

saveWidget(m, file="02 EDA/01 Map By SA4/Loss Of IncomeClaimsSA4.html")
m

```

## Map by State

Grouping by state highlights the clear difference of ACT compared to the other states. However, the state feature may not account for differences within a state 
```{r}
# prepare colour code
pal <- colorNumeric(
  palette = "BuPu", 
  domain = by_state$mean_prop_grossincurred,
  na.color = NA)

# prepare tooltip
mytext <- paste(
   "state: ", by_state$state, "<br/>", 
   "Number of policies: ", by_state$n_policies, "<br/>", 
   "Mean Number of Property claims: ", by_state$mean_prop_claims, "<br/>", 
   "Mean Log Gross Property Gross Incurred: ", by_state$mean_prop_grossincurred, sep="") %>%
  lapply(htmltools::HTML)

# create map
m <- leaflet(data = by_state) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lat=-27.833, lng=120 , zoom=4) %>%
  addCircles(
    lng = ~ long,
    lat = ~ lat,
    radius = ~mean_prop_claims*1500000,
    color = ~pal(mean_prop_grossincurred),
    weight = 1,
    popup = mytext,
    opacity = 0.8
  ) %>%
  addLegend("bottomleft",
            pal = pal, values = ~ mean_prop_grossincurred,
            title = "Mean Property Log Gross Incurred")

saveWidget(m, file="02 EDA/03 Map By Postcode/PropertyClaimsstate.html")
m
```

```{r}
# prepare colour code
pal <- colorNumeric(
  palette = "BuPu", 
  domain = by_state$mean_lossofinc_grossincurred,
  na.color = NA)

# prepare tooltip
mytext <- paste(
   "state: ", by_state$state, "<br/>", 
   "Number of policies: ", by_state$n_policies, "<br/>", 
   "Mean Number of Loss of Income claims: ", by_state$mean_lossofinc_claims, "<br/>", 
   "Mean Log Gross Loss of Income Gross Incurred: ", by_state$mean_lossofinc_grossincurred, sep="") %>%
  lapply(htmltools::HTML)

# create map
m <- leaflet(data = by_state) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lat=-27.833, lng=120 , zoom=4) %>%
  addCircles(
    lng = ~ long,
    lat = ~ lat,
    radius = ~mean_lossofinc_claims*50000000,
    color = ~pal(mean_lossofinc_grossincurred),
    weight = 1,
    popup = mytext,
    opacity = 0.8
  ) %>%
  addLegend("bottomleft",
            pal = pal, values = ~ mean_lossofinc_grossincurred,
            title = "Mean Loss of Income Log Gross Incurred")

saveWidget(m, file="02 EDA/02 Map By State/Loss of IncomeClaimsstate.html")
m
```

## Map by postcode

Postcode has the most information but suffers from high cardinality and sparsity issues due to low sample size in many of the postcodes

```{r}
# prepare colour code
pal <- colorNumeric(
  palette = "BuPu", 
  domain = by_postcode$mean_prop_grossincurred,
  na.color = NA)

# prepare tooltip
mytext <- paste(
   "postcode: ", by_postcode$riskpostcode, "<br/>", 
   "Number of policies: ", by_postcode$n_policies, "<br/>", 
   "Mean Number of Loss of Income claims: ", by_postcode$mean_prop_claims, "<br/>", 
   "Mean Log Gross Loss of Income Gross Incurred: ", by_postcode$mean_prop_grossincurred, sep="") %>%
  lapply(htmltools::HTML)

# create map
m <- leaflet(data = by_postcode) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lat=-27.833, lng=120 , zoom=4) %>%
  addCircleMarkers(
    lng = ~ long,
    lat = ~ lat,
    radius = ~sqrt(mean_prop_claims)*20,
    color = ~pal(mean_prop_grossincurred),
    weight = 1,
    popup = mytext,
    opacity = 0.8
  ) %>%
  addLegend("bottomleft",
            pal = pal, values = ~ mean_prop_grossincurred,
            title = "Mean Loss of Income Log Gross Incurred")

saveWidget(m, file="02 EDA/03 Map By Postcode/PropertyClaimspostcode.html")
m
```

```{r}
# prepare colour code
pal <- colorNumeric(
  palette = "BuPu", 
  domain = by_postcode$mean_lossofinc_grossincurred,
  na.color = NA)

# prepare tooltip
mytext <- paste(
   "postcode: ", by_postcode$riskpostcode, "<br/>", 
   "Number of policies: ", by_postcode$n_policies, "<br/>", 
   "Mean Number of Loss of Income claims: ", by_postcode$mean_lossofinc_claims, "<br/>", 
   "Mean Log Gross Loss of Income Gross Incurred: ", by_postcode$mean_lossofinc_grossincurred, sep="") %>%
  lapply(htmltools::HTML)

# create map
m <- leaflet(data = by_postcode) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  setView(lat=-27.833, lng=120 , zoom=4) %>%
  addCircleMarkers(
    lng = ~ long,
    lat = ~ lat,
    radius = ~sqrt(mean_lossofinc_claims)*60,
    color = ~pal(mean_lossofinc_grossincurred),
    weight = 1,
    popup = mytext,
    opacity = 0.8
  ) %>%
  addLegend("bottomleft",
            pal = pal, values = ~ mean_lossofinc_grossincurred,
            title = "Mean Loss of Income Log Gross Incurred")

saveWidget(m, file="02 EDA/03 Map By Postcode/ Loss of IncomeClaimspostcode.html")
m
```

# Sum insured

One of the clearest relationship was where as sum insured increases, the claim cost tends to also increase with this relationship being more pronounced for BI claims.
By looking at the higher distribution of SI for higher claim counts, we observe that policies with higher sum insured also have a higher claim frequency.

```{r}
myColors <- brewer.pal(nlevels(as.factor(policy_claims$claimcount_prop)), "BuPu")
names(myColors) <- levels(policy_claims$claimcount_prop)

##png(file="02 EDA/04 Sum Insured/Sum Insured vs Prop Claim Cost.##png", width=600, height=350)
policy_claims %>% 
  mutate(    
    claimcount_prop = as.factor(claimcount_prop),
    claimcount_lossofinc = as.factor(claimcount_lossofinc)) %>%
  filter(grossincurred_prop> 0) %>% 
  filter(suminsured_prop > 0) %>%
  ggplot(aes(x = log(suminsured_prop), y = log(grossincurred_prop) / as.numeric(claimcount_prop))) +
  geom_point(aes(col = claimcount_prop)) +
  scale_colour_manual(name = "Claim Count", values = myColors, drop = FALSE) +
  geom_smooth(method = "lm", color = "midnightblue")+
  xlab("Log Sum Insured") +
  ylab("Mean Log Claim Cost") +
  ggtitle("Sum Insured vs Property Claim Cost") +
  theme(legend.position="bottom")
##dev.off

##png(file="02 EDA/04 Sum Insured/Sum Insured vs Prop Claim Count.##png", width=600, height=350)
policy_claims %>%
  mutate(    
    claimcount_prop = as.factor(claimcount_prop),
    claimcount_lossofinc = as.factor(claimcount_lossofinc)) %>%  
  filter(suminsured_prop > 0) %>%
  ggplot(aes(x = claimcount_prop, y = log(suminsured_prop))) +
  geom_violin(aes(fill = claimcount_prop), draw_quantiles = 0.5) +
  scale_fill_manual(name = "Claim Count", values = myColors, drop = FALSE) +
  xlab("Claim Count") +
  ylab("Log Sum Insured") + 
  ggtitle("Sum Insured vs Property Claim Count") +
  theme(legend.position="bottom")
##dev.off
```

```{r}
myColors <- brewer.pal(nlevels(as.factor(policy_claims$claimcount_lossofinc)), "BuPu")
names(myColors) <- levels(policy_claims$claimcount_lossofinc)

##png(file="02 EDA/04 Sum Insured/Sum Insured vs Loss of Income Claim Count.##png", width=600, height=350)
policy_claims %>% 
  mutate(    
  claimcount_prop = as.factor(claimcount_prop),
  claimcount_lossofinc = as.factor(claimcount_lossofinc)) %>%
  filter(grossincurred_lossofinc > 0) %>% 
  filter(suminsured_lossofinc > 0) %>%
  ggplot(aes(x = log(suminsured_lossofinc), y = log(grossincurred_lossofinc / as.numeric(claimcount_lossofinc)))) +
  geom_point(aes(col = claimcount_lossofinc)) +
  scale_colour_manual(name = "Claim Count", values = myColors, drop = FALSE) +
  geom_smooth(method = "lm", color = "midnightblue")+
  xlab("Log Sum Insured") +
  ggtitle("Sum Insured vs Loss of Income Claim Cost") +
  ylab("Mean Log Claim Cost") + 
  theme(legend.position="bottom")
##dev.off

##png(file="02 EDA/04 Sum Insured/Sum Insured vs Loss of Income Claim Cost.##png", width=600, height=350)
policy_claims %>%
    mutate(    
    claimcount_prop = as.factor(claimcount_prop),
    claimcount_lossofinc = as.factor(claimcount_lossofinc)) %>%
  filter(suminsured_lossofinc > 0) %>%
  ggplot(aes(x = claimcount_lossofinc, y = log(suminsured_lossofinc))) +
  geom_violin(aes(fill = claimcount_lossofinc), draw_quantiles = 0.5) +
  scale_fill_manual(name = "Claim Count", values = myColors, drop = FALSE) +
  xlab("Claim Count") +
  ylab("Log Sum Insured")+ 
  ggtitle("Sum Insured vs Loss of Income Claim Count") +
  theme(legend.position="bottom")
##dev.off
```

```{r}
onlynumeric <- policy_claims %>%
  dplyr:: select(
    frequency_prop,
    severity_prop,
    frequency_bi,
    severity_bi
  )

corr = cor(onlynumeric, method = "pearson")

#png(file="02 EDA/05 Correlation/Corrplot.#png", width=1000, height=600)

corrplot(corr, method="color",   
         col.lim = c(0,1), 
         col=brewer.pal(n=10, name="BuPu"),
         type="lower", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )

#dev.off()
```

# Property Claims - Building Characteristics

Further we can observe different claim experiences for the different building materials with the bars reflecting severity and the line representing frequency. 

Including these variables may help predictive power as there appears to be strong differences in claim experience between levels within the variable. For example, policies with Sprinkler 2 are more likely to make claims that are more frequent and of greater severity.

## Building Age

```{r}
building_age <- policy_claims %>%
  group_by(building_age) %>%
  summarise(
    n_obs = n(),
    avg_prop_claim = mean(grossincurred_prop / claimcount_prop, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_prop_no = mean(claimcount_prop),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/06 Property - Building Characteristics/Building_Age.#png", width=600, height= 550)
coeff <- max(building_age$avg_prop_no) / max(building_age$avg_prop_claim)

avg_prop_noColor <- col4
avg_prop_claimColor <- col1

ggplot(building_age, aes(x=building_age)) +
  
  geom_bar( aes(y=avg_prop_claim), stat="identity", size=.1, fill=avg_prop_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_prop_no / coeff), size=2, color=avg_prop_noColor, group = 1) +
  geom_text(aes(y=avg_prop_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 2.5) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Building Age") +
  theme(
    axis.title.y = element_text(color = avg_prop_noColor),
    axis.title.y.right = element_text(color = avg_prop_claimColor)
  ) 
#dev.off
```

## Building Type

```{r}
building_type <- policy_claims %>%
  group_by(building_type) %>%
  summarise(
    n_obs = n(),
    avg_prop_claim = mean(grossincurred_prop / claimcount_prop, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_prop_no = mean(claimcount_prop),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/06 Property - Building Characteristics/Building_Type.#png", width=600, height= 550)
coeff <- max(building_type$avg_prop_no) / max(building_type$avg_prop_claim)

avg_prop_noColor <- col4
avg_prop_claimColor <- col1

ggplot(building_type, aes(x=building_type)) +
  geom_bar( aes(y=avg_prop_claim), stat="identity", size=.1, fill=avg_prop_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_prop_no / coeff), size=2, color=avg_prop_noColor, group = 1) +
  geom_text(aes(y=avg_prop_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 2.5) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Building Type") +
  theme(
    axis.title.y = element_text(color = avg_prop_noColor),
    axis.title.y.right = element_text(color = avg_prop_claimColor)
  ) 
#dev.off
```
## Construction Walls
```{r}
construction_walls <- policy_claims %>%
  group_by(construction_walls) %>%
  summarise(
    n_obs = n(),
    avg_prop_claim = mean(grossincurred_prop / claimcount_prop, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_prop_no = mean(claimcount_prop),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/06 Property - Building Characteristics/Construction_Walls.#png", width=600, height= 550)

coeff <- max(construction_walls$avg_prop_no) / max(construction_walls$avg_prop_claim)

avg_prop_noColor <- col4
avg_prop_claimColor <- col1

ggplot(construction_walls, aes(x=construction_walls)) +
  
  geom_bar( aes(y=avg_prop_claim), stat="identity", size=.1, fill=avg_prop_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_prop_no / coeff), size=2, color=avg_prop_noColor, group = 1) +
  geom_text(aes(y=avg_prop_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 3) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Construction Walls") +
  theme(
    axis.title.y = element_text(color = avg_prop_noColor),
    axis.title.y.right = element_text(color = avg_prop_claimColor)
  ) 
#dev.off
```

## Construction Floor

```{r}
construction_floor <- policy_claims %>%
  group_by(construction_floor) %>%
  summarise(
    n_obs = n(),
    avg_prop_claim = mean(grossincurred_prop / claimcount_prop, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_prop_no = mean(claimcount_prop),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/06 Property - Building Characteristics/Construction_Floor.#png", width=600, height= 550)

coeff <- max(construction_floor$avg_prop_no) / max(construction_floor$avg_prop_claim)

avg_prop_noColor <- col4
avg_prop_claimColor <- col1

ggplot(construction_floor, aes(x=construction_floor)) +
  
  geom_bar( aes(y=avg_prop_claim), stat="identity", size=.1, fill=avg_prop_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_prop_no / coeff), size=2, color=avg_prop_noColor, group = 1) +
  geom_text(aes(y=avg_prop_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 3) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Construction Floor") +
  theme(
    axis.title.y = element_text(color = avg_prop_noColor),
    axis.title.y.right = element_text(color = avg_prop_claimColor)
  ) 
#dev.off
```

## Sprinkler Type

```{r}
sprinkler_type <- policy_claims %>%
  group_by(sprinkler_type) %>%
  summarise(
    n_obs = n(),
    avg_prop_claim = mean(grossincurred_prop / claimcount_prop, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_prop_no = mean(claimcount_prop),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/06 Property - Building Characteristics/Sprinkler.#png", width=600, height= 550)

coeff <- max(sprinkler_type$avg_prop_no) / max(sprinkler_type$avg_prop_claim)

avg_prop_noColor <- col4
avg_prop_claimColor <- col1

ggplot(sprinkler_type, aes(x=sprinkler_type)) +
  geom_bar( aes(y=avg_prop_claim), stat="identity", size=.1, fill=avg_prop_claimColor, color="black", alpha=.7) +
  geom_text(aes(y=avg_prop_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 4) +
  geom_line(aes(y=avg_prop_no / coeff), size=2, color=avg_prop_noColor, group = 1) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Sprinkler Type") +
  theme(
    axis.title.y = element_text(color = avg_prop_noColor),
    axis.title.y.right = element_text(color = avg_prop_claimColor)
  ) 
#dev.off
```
# Loss Of Income Claims - Building Characteristics

## Building Age

```{r}
building_age <- policy_claims %>%
  group_by(building_age) %>%
  summarise(
    n_obs = n(),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_no = mean(claimcount_lossofinc),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )
#png(file="02 EDA/07 Loss of Income - Building Characteristics/Building_Age.#png", width=600, height= 550)
coeff <- max(building_age$avg_lossofinc_no) / max(building_age$avg_lossofinc_claim)

avg_lossofinc_noColor <- col4
avg_lossofinc_claimColor <- col3

ggplot(building_age, aes(x=building_age)) +
  
  geom_bar( aes(y=avg_lossofinc_claim), stat="identity", size=.1, fill=avg_lossofinc_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_lossofinc_no / coeff), size=2, color=avg_lossofinc_noColor, group = 1) +
  geom_text(aes(y=avg_lossofinc_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 2.5) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Building Age") +
  theme(
    axis.title.y = element_text(color = avg_lossofinc_noColor),
    axis.title.y.right = element_text(color = avg_lossofinc_claimColor)
  ) 
#dev.off
```

## Building Type

```{r}
building_type <- policy_claims %>%
  group_by(building_type) %>%
  summarise(
    n_obs = n(),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_no = mean(claimcount_lossofinc),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/07 Loss of Income - Building Characteristics/Building_Type.#png", width=600, height= 550)

coeff <- max(building_type$avg_lossofinc_no) / max(building_type$avg_lossofinc_claim)

avg_lossofinc_noColor <- col4
avg_lossofinc_claimColor <- col3

ggplot(building_type, aes(x=building_type)) +
  geom_bar( aes(y=avg_lossofinc_claim), stat="identity", size=.1, fill=avg_lossofinc_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_lossofinc_no / coeff), size=2, color=avg_lossofinc_noColor, group = 1) +
  geom_text(aes(y=avg_lossofinc_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 2.5) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Building Type") +
  theme(
    axis.title.y = element_text(color = avg_lossofinc_noColor),
    axis.title.y.right = element_text(color = avg_lossofinc_claimColor)
  ) 
#dev.off
```

## Construction Walls

```{r}
construction_walls <- policy_claims %>%
  group_by(construction_walls) %>%
  summarise(
    n_obs = n(),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_no = mean(claimcount_lossofinc),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/07 Loss of Income - Building Characteristics/Construction_Walls.#png", width=600, height= 550)

coeff <- max(construction_walls$avg_lossofinc_no) / max(construction_walls$avg_lossofinc_claim)

avg_lossofinc_noColor <- col4
avg_lossofinc_claimColor <- col3

ggplot(construction_walls, aes(x=construction_walls)) +
  
  geom_bar( aes(y=avg_lossofinc_claim), stat="identity", size=.1, fill=avg_lossofinc_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_lossofinc_no / coeff), size=2, color=avg_lossofinc_noColor, group = 1) +
  geom_text(aes(y=avg_lossofinc_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 3) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Construction Walls") +
  theme(
    axis.title.y = element_text(color = avg_lossofinc_noColor),
    axis.title.y.right = element_text(color = avg_lossofinc_claimColor)
  ) 
```

## Construction Floor

```{r}
construction_floor <- policy_claims %>%
  group_by(construction_floor) %>%
  summarise(
    n_obs = n(),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_no = mean(claimcount_lossofinc),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )

#png(file="02 EDA/07 Loss of Income - Building Characteristics/Construction_Floor.#png", width=600, height= 550)
coeff <- max(construction_floor$avg_lossofinc_no) / max(construction_floor$avg_lossofinc_claim)

avg_lossofinc_noColor <- col4
avg_lossofinc_claimColor <- col3

ggplot(construction_floor, aes(x=construction_floor)) +
  
  geom_bar( aes(y=avg_lossofinc_claim), stat="identity", size=.1, fill=avg_lossofinc_claimColor, color="black", alpha=.7) + 
  geom_line(aes(y=avg_lossofinc_no / coeff), size=2, color=avg_lossofinc_noColor, group = 1) +
  geom_text(aes(y=avg_lossofinc_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 3) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Construction Floor") +
  theme(
    axis.title.y = element_text(color = avg_lossofinc_noColor),
    axis.title.y.right = element_text(color = avg_lossofinc_claimColor)
  ) 
#dev.off
```

## Sprinkler Type

```{r}
sprinkler_type <- policy_claims %>%
  group_by(sprinkler_type) %>%
  summarise(
    n_obs = n(),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_claim = mean(grossincurred_lossofinc / claimcount_lossofinc, na.rm = TRUE),
    avg_lossofinc_no = mean(claimcount_lossofinc),
    avg_lossofinc_no = mean(claimcount_lossofinc)
  )
#png(file="02 EDA/07 Loss of Income - Building Characteristics/Sprinkler.#png", width=600, height= 550)
coeff <- max(sprinkler_type$avg_lossofinc_no) / max(sprinkler_type$avg_lossofinc_claim)

avg_lossofinc_noColor <- col4
avg_lossofinc_claimColor <- col3

ggplot(sprinkler_type, aes(x=sprinkler_type)) +
  geom_bar( aes(y=avg_lossofinc_claim), stat="identity", size=.1, fill=avg_lossofinc_claimColor, color="black", alpha=.7) +
  geom_text(aes(y=avg_lossofinc_claim, label = paste("n_obs: ", format(n_obs, nsmall = 2))), vjust = 1.5, colour = "white", size = 4) +
  geom_line(aes(y=avg_lossofinc_no / coeff), size=2, color=avg_lossofinc_noColor, group = 1) +
  scale_y_continuous(
    name = "Mean Number of Claims",
    sec.axis = sec_axis(~.*coeff, name="Mean Claim Size")
  ) + 
  ggtitle("Sprinkler Type") +
  theme(
    axis.title.y = element_text(color = avg_lossofinc_noColor),
    axis.title.y.right = element_text(color = avg_lossofinc_claimColor)
  ) 

#dev.off
```

# Occupation

For property claims, it can be seen that Gas and Water has significantly poorer claims experience.

```{r}
occupations <- sort(as.character(unique(policy_claims$occupation_risk)))
temp <- occupations[17]
occupations[17] <- occupations[18]
occupations[18] <- temp
reorg_occupation <- policy_claims %>%
  dplyr:: mutate(
    occupation_risk = factor(occupation_risk, occupations),
    mean = as.factor(round(ave(log_grossincurred_prop, as.factor(occupation_risk), FUN=mean)))
  )
##png(file="02 EDA/08 Occupations/PropertyOccs.#png", width=1200, height= 550)
reorg_occupation%>%
  group_by(occupation_risk)%>%
  mutate(mean=mean(log_grossincurred_prop, na.rm=TRUE))%>%
  ggplot(aes(factor(occupation_risk),log_grossincurred_prop))+
  geom_boxplot(aes(fill=mean)) + 
  scale_fill_gradientn(colors = brewer.pal(9, "BuPu")) +
  xlab("Occupation Risk") +
  ylab("Average Log Gross Incurred") +
  ggtitle("Property - Occupation Risk") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) 
##dev.off()
```

Due to how the nature of an industry operates would affect business intnerruption claims, industries such as C - Manufacturing, G - Retail and L Rental for Business Services have higher average gross incurred

```{r}
occupations <- sort(as.character(unique(policy_claims$occupation_risk)))
temp <- occupations[17]
occupations[17] <- occupations[18]
occupations[18] <- temp
reorg_occupation <- policy_claims %>%
  dplyr:: mutate(
    occupation_risk = factor(occupation_risk, occupations),
    mean = as.factor(round(ave(log_grossincurred_lossofinc, as.factor(occupation_risk), FUN=mean)))
  )

##png(file="02 EDA/08 Occupations/LossofincOccs.#png", width=1200, height= 550)
reorg_occupation%>%
  group_by(occupation_risk)%>%
  mutate(mean=mean(log_grossincurred_lossofinc, na.rm=TRUE))%>%
  ggplot(aes(factor(occupation_risk),log_grossincurred_lossofinc))+
  geom_boxplot(aes(fill=mean)) + 
  scale_fill_gradientn(colors = brewer.pal(9, "BuPu")) +
  xlab("Occupation Risk") +
  ylab("Average Log Gross Incurred") +
  ggtitle("Loss of Income - Occupation Risk") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) 
##dev.off()
```

# Interaction

```{r}
ass_rfct %>%
  summarise(both = sum(grossincurred_prop > 0 & grossincurred_lossofinc > 0, na.rm = TRUE),
            only_prop = sum(grossincurred_prop > 0 & grossincurred_lossofinc == 0, na.rm = TRUE),
            only_lossofinc = sum(grossincurred_prop == 0 & grossincurred_lossofinc >0, na.rm = TRUE),
            no_claim = sum(grossincurred_prop == 0 & grossincurred_lossofinc == 0, na.rm = TRUE))
```

## Property Claim Cost vs BI Claim Cost

Clear positive trend between Property and BI claim cost.

```{r}
myColors <- brewer.pal(nlevels(as.factor(policy_claims$indem_per_grp)), "BuPu")
names(myColors) <- levels(policy_claims$indem_per_grp)

##png(file="02 EDA/09 Interaction/GrossIncurred.#png", width=600, height= 500)
policy_claims %>%
  ggplot(aes(x = log_grossincurred_prop / claimcount_prop, y = log_grossincurred_lossofinc / claimcount_lossofinc)) +
  geom_point(aes(col = indem_per_grp))+
  scale_colour_manual(name = "Indemnity period (BI)", values = myColors, drop = FALSE) +
  xlab("Mean Log Claim Cost (Prop)") +
  ylab("Mean Log Claim Cost (BI)") +
  ggtitle("Property Claim Cost vs BI Claim Cost for Policies That Incurred Both Losses") +
  geom_smooth(method = glm, col = col1)
##dev.off()
```

## Distribution of polciies with BI cover vs without

```{r}
policy_claims %>%
  ggplot(aes(x = LossofIncome_cover, y = log_grossincurred_prop)) +
  geom_boxplot()
```

## Exploring Claims Made Simultaneously vs Rest

By comparing gross incurred for policies that make both claims vs the others, we see that when both claims are made in conjunction, a policy tends to make larger claims

```{r}
claim_type <- ass_rfct %>%
  mutate(
    claim_type = case_when(
        grossincurred_prop > 0 & grossincurred_lossofinc > 0 ~ "Both Prop and BI Claim",
        grossincurred_prop > 0 & grossincurred_lossofinc == 0 ~ "Only Prop Claim",
        grossincurred_prop == 0 & grossincurred_lossofinc >0 ~ "Only BI Claim",
        grossincurred_prop == 0 & grossincurred_lossofinc == 0 ~ "No Claim",
        is.na(grossincurred_lossofinc) & grossincurred_prop == 0 ~ "No Claim",
        is.na(grossincurred_lossofinc) & grossincurred_prop > 0 ~ "Only Prop Claim"),
    claim_type = factor(claim_type, levels=c("No Claim",  "Both Prop and BI Claim", "Yeet", "Only Prop Claim","Only BI Claim"))
  )

myColors <- brewer.pal(nlevels(as.factor(claim_type$claim_type)), "BuPu")
names(myColors) <- levels(claim_type$claim_type)

#png(file="02 EDA/09 Interaction/Claim Type - Property.#png", width=600, height= 500)
claim_type %>%
  filter(grossincurred_prop > 0) %>%
  ggplot(aes(x = log(grossincurred_prop), col = claim_type, fill = claim_type)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(name = "Claim Type", values = myColors)+
  scale_colour_manual(name = "Claim Type", values = myColors) +
  xlab("Log Property Claim Size")

#dev.off()

#png(file="02 EDA/09 Interaction/Claim Type - BI.#png", width=600, height= 500)
claim_type %>%
  filter(grossincurred_lossofinc > 0) %>%
  ggplot(aes(x = log(grossincurred_lossofinc), col = claim_type, fill = claim_type)) +
  scale_fill_manual(name = "Claim Type", values = myColors)+
  scale_colour_manual(name = "Claim Type", values = myColors)+
  geom_density(alpha = 0.5) +
  xlab("Log BI Claim Size")

#dev.off()
```


