---
title: "Insurancerating Package Demonstration"
output: html_notebook
---

This Notebook demonstrates the use of the `insurancerating` package. The package helps to implement GLMs by executing all steps needed to construct a risk premium from the data. 

It provides a data driven strategy for the construction of tariff classes in P&C insurance. 

The goal is to **bin the continuous factors**, to then produce categorical risk factors which capture the effect of the covariate on the response in an accurate way, while being easy to use in a generalized linear model (GLM).


```{r setup, include = FALSE}
proj_dir = rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir = proj_dir)
```

```{r load_envr, include=FALSE}
source("00 source.R")
library(insurancerating)
load("00 envr/Compulsory/policy_claims.Rda")
```

Insurancerating requires each row to be each policy. Policy claims already fulfils this requirement. Furthermore, the dataset `policy_claims` is primarily categorical.

---

# Policy Claims Trial Run


## Full Data Set

Have both LoI and Proper

### Property Claims

```{r}
full_dat <- policy_claims%>%
  select(-contains("lossofinc"))

cat_variables <- colnames(full_dat)[c(8:15, 17)]

univ_full <- 
  tibble(var_analysis = cat_variables)%>%
  mutate(univ_df = map(
    var_analysis, 
    ~univariate(full_dat,
           x = vec_ext(.), 
           nclaims = claimcount_prop,
           exposure = exposure,
           severity = grossincurred_prop)))
```

#### Geocode

There are 6 areas to look out for, especially those with relatively largge exposures

```{r}
univ_full$univ_df[[1]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)

univ_full$univ_df[[1]]%>%
  as_tibble()%>%
  arrange(desc(average_severity))%>%
  filter(average_severity > 50000 | risk_premium > 50000, exposure > 5)%>%
  print()
```

#### Occupation

```{r}
univ_full$univ_df[[8]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Building age

```{r}
univ_full$univ_df[[3]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Building type 

```{r}
univ_full$univ_df[[4]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### State

```{r}
univ_full$univ_df[[2]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### Construction Walls

```{r}
univ_full$univ_df[[5]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### Construction floor

```{r}
univ_full$univ_df[[6]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Indenmity

```{r}
univ_full$univ_df[[9]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

## Only Property Coverage

### Property Claims

```{r}
prop_dat <- policy_claims%>%
  filter(!LossofIncome_cover)%>%
  select(-contains("lossofinc"))

prop_cat_variables <- colnames(prop_dat)[c(8:15, 17)]

univ_prop <- 
  tibble(var_analysis = prop_cat_variables)%>%
  mutate(univ_df = map(
    var_analysis, 
    ~univariate(prop_dat,
           x = vec_ext(.), 
           nclaims = claimcount_prop,
           exposure = exposure,
           severity = grossincurred_prop)))
```

#### Geocode

There are 6 areas to look out for, especially those with relatively largge exposures

```{r}
univ_prop$univ_df[[1]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)

univ_prop$univ_df[[1]]%>%
  as_tibble()%>%
  arrange(desc(average_severity))%>%
  filter(average_severity > 50000 | risk_premium > 50000, exposure > 5)%>%
  print()
```

#### Occupation

```{r}
univ_prop$univ_df[[8]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Building age

```{r}
univ_prop$univ_df[[3]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Building type 

```{r}
univ_prop$univ_df[[4]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### State

```{r}
univ_prop$univ_df[[2]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### Construction Walls

```{r}
univ_prop$univ_df[[5]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### Construction floor

```{r}
univ_prop$univ_df[[6]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```


### Sum Insured Treatment

```{r}
prop_sumins_freq <- fit_gam(
  data = prop_dat,
  nclaims = claimcount_prop,
  x = suminsured_prop,
  exposure = exposure
)

prop_sumins_severity <- fit_gam(
  data = prop_dat,
  nclaims = claimcount_prop,
  x = suminsured_prop,
  exposure = exposure,
  amount = grossincurred_prop,
  model = "severity"
)
```

```{r}
autoplot(prop_sumins_freq, show_observations = T, remove_outliers = 40)
autoplot(prop_sumins_severity, show_observations = T, remove_outliers=1000000)
```

## Property + LoI Coverage

### Property Claims

```{r}
prop_loi_dat <- policy_claims%>%
  filter(LossofIncome_cover)%>%
  select(-contains("lossofinc"))

prop_loi_cat_variables <- colnames(prop_loi_dat)[c(8:15, 17)]

univ_prop_loi <- 
  tibble(var_analysis = prop_loi_cat_variables)%>%
  mutate(univ_df = map(
    var_analysis, 
    ~univariate(prop_loi_dat,
           x = vec_ext(.), 
           nclaims = claimcount_prop,
           exposure = exposure,
           severity = grossincurred_prop)))
```

#### Geocode

There are 6 areas to look out for, especially those with relatively largge exposures

```{r}
univ_prop_loi$univ_df[[1]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)

univ_prop_loi$univ_df[[1]]%>%
  as_tibble()%>%
  arrange(desc(average_severity))%>%
  filter(average_severity > 50000 | risk_premium > 50000, exposure > 5)%>%
  print()
```

#### Occupation

```{r}
univ_prop_loi$univ_df[[8]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Building age

```{r}
univ_prop_loi$univ_df[[3]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Building type 

```{r}
univ_prop_loi$univ_df[[4]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### State

```{r}
univ_prop_loi$univ_df[[2]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### Construction Walls

```{r}
univ_prop_loi$univ_df[[5]]%>%
  autoplot(., show_plots = c(1, 2), background = T)
```

#### Construction floor

```{r}
univ_prop_loi$univ_df[[6]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

#### Indenmity

```{r}
univ_prop_loi$univ_df[[9]]%>%
  autoplot(., show_plots = c(1, 2, 3), background = T)
```

