---
title: "Exploratory Data Analysis"
output: html_notebook
---

This R Notebook records the exploratory analysis to fulfil the part 1 requirements of the ACTL4305 Assignment

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
theme_set(theme_classic())
options(scipen=999)
options(digit=3)
```


```{r, include=FALSE}
source("00 source.R")
load("00 envr/Compulsory/ass_copy.R")
load("00 envr/Compulsory/policy_df.R")
load("00 envr/Compulsory/policy_history_df.R")
```

# On a property historical basis (ignoring policy instance)
## Define variable types
```{r}
num_vars <- names(policy_history_df)[sapply(policy_history_df, is.numeric)]
logical_vars <- names(policy_history_df)[sapply(policy_history_df, is.logical)]
factor_vars <- names(policy_history_df)[sapply(policy_history_df, is.factor)]
date_vars <- names(policy_history_df)[sapply(policy_history_df, is.Date)]
```

## Number of claims
```{r}
policy_history_df %>%
  ggplot(aes(n_claims_prop)) +
  geom_bar() + 
  ggtitle("Number of property claims per property") +
  labs(y = "Count", x = "Number of property claims")

policy_history_df %>%
  filter(n_claims_prop > 0) %>%
  ggplot(aes(n_claims_prop)) +
  geom_bar() +
  ggtitle("Number of property claims per property excluding 0 claims") +
  labs(y = "Count", x = "Number of property claims")

policy_history_df %>%
  ggplot(aes(n_claims_lossofinc)) +
  geom_bar() +
  ggtitle("Number of loss of income claims per property") +
  labs(y = "Count", x = "Number of loss of income claims")

policy_history_df %>%  
  filter(n_claims_lossofinc > 0) %>%
  ggplot(aes(n_claims_lossofinc)) +
  geom_bar() + 
  ggtitle("Number of loss of income claims per property excluding 0 claims") +
  labs(y = "Count", x = "Number of loss of income claims")
```
## Gross incurred

```{r}
policy_history_df %>%
  filter(total_grossincurred_lossofinc > 0) %>%
  ggplot(aes(total_grossincurred_lossofinc)) +
  geom_density()

policy_history_df %>%
  filter(total_grossincurred_lossofinc > 0) %>%
  group_by(total_grossincurred_lossofinc) %>%
  count()

policy_history_df %>%
  filter(total_grossincurred_prop > 0) %>%
  ggplot(aes(total_grossincurred_prop)) +
  geom_density()

policy_history_df %>%
  filter(total_grossincurred_prop > 0) %>%
  group_by(total_grossincurred_prop) %>%
  count()
```

## Sum Insured

```{r}
policy_history_df %>%
  ggplot(aes(median_SI_lossofinc)) +
  geom_density()

policy_history_df %>%
  group_by(median_SI_lossofinc) %>%
  count()

policy_history_df %>%
  ggplot(aes(median_SI_prop)) +
  geom_density()

policy_history_df %>%
  group_by(median_SI_prop) %>%
  count()
```


## SI violin plot
```{r}
plot_data <- policy_history_df %>%
  mutate(claim_prop = as.factor(ifelse(n_claims_prop > 0, TRUE, FALSE)),
         claim_lossofinc = as.factor(ifelse(n_claims_lossofinc > 0, TRUE, FALSE)),
         n_claims_prop = as.factor(n_claims_prop),
         n_claims_lossofinc = as.factor(n_claims_lossofinc)
         )

plot_data %>%
  ggplot(aes(x = n_claims_prop, y = median_SI_prop, fill = n_claims_prop)) +
  geom_violin(draw_quantiles = 0.5) + 
  ylim(0, 5000000)+
  ggtitle("Property Cover - Sum Insured vs Number of Claims") +
  labs(y = "Sum insured", x = "Number of property claims")

plot_data %>%
  ggplot(aes(x = claim_prop, y = median_SI_prop, fill = claim_prop)) +
  geom_violin(draw_quantiles = 0.5) + 
  ylim(0, 5000000)+
  ggtitle("Property Cover - Sum Insured vs Claim Filed") +
  labs(y = "Sum insured", x = "Property Claim Filed")

plot_data %>%
  ggplot(aes(x = n_claims_lossofinc, y = median_SI_lossofinc, fill = n_claims_lossofinc)) +
  geom_violin(draw_quantiles = 0.5) +
  ylim(0, 1000000)+
  ggtitle("Loss of Income - Sum Insured vs Number of Claims") +
  labs(y = "Sum insured", x = "Number of Loss of Income claims")

plot_data %>%
  ggplot(aes(x = claim_lossofinc, y = median_SI_lossofinc, fill = claim_lossofinc)) +
  geom_violin(draw_quantiles = 0.5)+
  ylim(0, 1000000)+
  ggtitle("Loss of Income - Sum Insured vs Claim Filed") +
  labs(y = "Sum insured", x = "Loss of Income Claim Filed")

```

## Numerical data histogram

```{r}

plot_data %>%
  ggplot(aes(median_SI_prop, fill = n_claims_prop, color = n_claims_prop)) +
  geom_density(alpha = 0.3)+ 
  xlim(0, 20000000)

plot_data %>%
  ggplot(aes(median_SI_lossofinc, fill = n_claims_lossofinc, color = n_claims_lossofinc)) +
  geom_density(alpha = 0.3)+ 
  xlim(0, 3000000)
```

```{r}
plot_data %>%
  ggplot(aes(median_SI_prop, avgincurred_prop, color = total_grossincurred_prop, size = total_grossincurred_prop)) +
  geom_point(alpha = 0.3)

plot_data %>%
  ggplot(aes(median_SI_prop, avgincurred_prop, color = total_grossincurred_prop, size = total_grossincurred_prop)) +
  geom_point(alpha = 0.3) +
  xlim(0,10000000) +
  ylim(0,200000)

plot_data %>%
  ggplot(aes(median_SI_lossofinc, avgincurred_lossofinc, color = total_grossincurred_lossofinc, size = total_grossincurred_lossofinc)) +
  geom_point(alpha = 0.3)

plot_data %>%
  ggplot(aes(median_SI_lossofinc, avgincurred_lossofinc, color = total_grossincurred_lossofinc, size = total_grossincurred_lossofinc)) +
  geom_point(alpha = 0.3) +
  xlim(0,10000000) +
  ylim(0,200000)
```

## Relationship between prop and loss of income claims
```{r}
plot_data %>%
  ggplot(aes(avgincurred_lossofinc, avgincurred_prop)) +
  geom_point(alpha = 0.3) + 
  xlim(0,100000) +
  ylim(0, 2000000)

plot_data %>%
  ggplot(aes(total_grossincurred_lossofinc, total_grossincurred_prop)) +
  geom_point(alpha = 0.3)

plot_data %>%
  filter(LossofIncome_cover = TRUE) %>%
  group_by(n_claims_prop, n_claims_lossofinc) %>%
  count(n_claims_prop, n_claims_lossofinc) %>%
	ggplot(mapping = aes(x = n_claims_prop, y = n_claims_lossofinc)) +
	geom_tile(mapping = aes(fill = n))


```

## Investigating level of cover vs SI and claim behaviour
```{r}
plot_data %>%
  ggplot(aes(median_SI_prop, avgincurred_prop, color = LossofIncome_cover, size = avgincurred_prop)) +
  geom_point(alpha = 0.3)


plot_data %>%
  ggplot(aes(median_SI_prop, fill = LossofIncome_cover, color = LossofIncome_cover)) +
  geom_boxplot()+
  xlim(0, 5000000)

plot_data %>%
  filter(avgincurred_prop > 0) %>%
  ggplot(aes(avgincurred_prop, fill = LossofIncome_cover, color = LossofIncome_cover)) +
  geom_boxplot()+ 
  xlim(0, 1000000)

plot_data %>%
  ggplot(aes(x = n_claims_prop, y = median_SI_prop, fill = n_claims_prop)) +
  geom_violin(draw_quantiles = 0.5) + 
  facet_grid(cols = vars(LossofIncome_cover))+
  ylim(0, 5000000)

```

# Categorical Variables (by Andrew)
```{r}
policy_df %>%
  ggplot(aes(state)) +
  geom_bar()
```

```{r}
policy_df %>%
  ggplot(aes(building_age)) +
  geom_bar()
```

```{r}
policy_df %>%
  ggplot(aes(building_type)) +
  geom_bar()
```

```{r}
policy_df %>%
  ggplot(aes(construction_walls)) +
  geom_bar()
```

```{r}
policy_df %>%
  ggplot(aes(construction_floor)) +
  geom_bar()
```

```{r}
policy_df %>%
  #filter(sprinkler_type != 0) %>%
  ggplot(aes(sprinkler_type)) +
  geom_bar()
```

```{r}
policy_df %>%
  ggplot(aes(occupation_risk)) +
  geom_bar()
```

